
\section{Introduction and Overview}

The aim of this project is to study the metal to Mott-Insulator phase transition exhibited by the Fermi-Hubbard model.

The model consists of a lattice with a single-level atom at every site. The electrons can only hop from a site to a nearest neighbor one, and only interact with each other if they are at the same site. The Hamiltonian of this model is given by:
%
\begin{equation}
\mathcal{H} = -t\sum_{<i,j>}c_{i,\sigma}^\dagger c_{j,\sigma} + h.c.+ U\sum_i n_{i,\uparrow}n_{i,\downarrow} + \mu\sum_i (n_{i,\uparrow} + n_{i,\downarrow})
\end{equation}
%
where $t$ is the hopping rate, $U$ the strenght of the interaction, which by varying will lead to the phase transition, and $\mu$ the chemical potential.

This model is studied here in the framework of Dynamical Mean Field Theory~(DMFT), and the quantity of interest is the local Green's function, given by:

\begin{equation}\label{orGreen's}
G_{\text{loc}} (\tau - \tau^\prime) = -\braket{Tc_{i\sigma}(\tau)c_{i\sigma}^\dagger(\tau^\prime)} 
\end{equation}
by means of which it will then be possible to compute the \emph{Spectral Function} as a measure of elementary excitation of the system.

The main idea behind the DMFT approach is similar in spirit to the classical mean field approximation and consists of solving the problem of a single atom coupled to a thermal bath and mapping this to our original lattice problem via a self-consistency relation.
Such single atom problem is described by the Hamiltonian of the so called \emph{Anderson Impurity Model} (AIM), given by:
%
\begin{equation}\label{AIMH}
\begin{array}{c}
\mathcal{H}_{\text{AIM}} = \mathcal{H}_{\text{atom}} + \mathcal{H}_{\text{bath}} + \mathcal{H}_{\text{coupling}}
\quad \quad
\text{with}
\quad
\mathcal{H}_{\text{atom}} = Un_\uparrow^cn_\downarrow^c - \mu (n_\uparrow^c+n_\downarrow^c),
\\ 
\\ 
\mathcal{H}_{\text{bath}} = \sum_{l,\sigma}\tilde{\epsilon}_la_{l\sigma}^\dagger a_{l\sigma},
\quad
\mathcal{H}_{\text{coupling}} = \sum_{l,\sigma}V_l(a_{l\sigma}^\dagger c_{\sigma} + c_{\sigma}^\dagger a_{l\sigma})
\, .
\end{array}
\end{equation}
%
Here, the $a_l$ operators describe the fermionic degrees of freedom of the bath, while the $\tilde{\epsilon_l}$ and the $V_l$ are parameters, which must been chosen appropriately (such that the impurity Green's function of \eqref{AIMH} coincides with the local lattice one). In section \ref{sec_impsolv} we will see how, after integrating out the bath, these parameters enter into an effective action for the singled out electron. Thereby, the impurity problem is defined with a given bare propagator $G_0$ and a value of the interaction parameter $U$.

At this point, the mean field approximation comes into play. First of all, we notice that the self-energy $\Sigma_\mathrm{imp}$ of the effective AIM with full Green's function $G$ is given by:
%
\begin{equation}\label{s_imp}
\Sigma_{\text{imp}}(i\omega_n) \equiv G_0^{-1}(i\omega_n) - G^{-1}(i\omega_n)
\end{equation}
%
Of course we can also consider the self-energy of our original lattice problem having a dispersion relation $\varepsilon_{\bold{k}}$, which gives the relation:
%
\begin{equation} \label{not_summed_over}
G_{\text{lattice}}(\bold{k},i\omega_n) = \frac{1}{i\omega_n - \varepsilon_{\bold{k}} + \mu -\Sigma_{\text{lattice}}(\bold{k},i\omega_n)}
\quad \text{with}
\quad
\varepsilon_\bold{k} \equiv t \sum_j e^{i\bold{k} \cdot ( \bold{R_i} - \bold{R_j} )}
.
\end{equation}


The approximation, now, consists of assuming that at each site the lattice self-energy coincides with the impurity self-energy and has its off-diagonal elements vanish:
%
\begin{equation}
\Sigma_{ii} \simeq \Sigma_{\text{imp}} ~ , \Sigma_{i\neq j} \simeq 0
\quad \Rightarrow
\quad
\Sigma_{\text{lattice}}(\bold{k},i\omega_n) = \Sigma_{\text{imp}}(i\omega_n)
.
\end{equation}
%
%\todo{look up this statement}This is a consistent approximation only given that it uniquely determines the local Green's function, which, by assumption, is the impurity problem Green's function. 
We average the momentum dependent Green's function \eqref{not_summed_over} over $\bold{k}$ to obtain the local propagator \eqref{orGreen's} and use \eqref{s_imp} to relate the impurity and lattice problem with each other:
%
\begin{equation} \label{Gloc_dos}
G_{\text{loc}}(i\omega_n) = \frac{1}{N} \sum_{\bold{k}} G(\bold{k},i\omega_n) 
 = \int \dif \varepsilon \frac{D(\varepsilon)}{i\omega_n - \varepsilon + \mu - \Sigma_{\text{imp}}(i\omega_n)},
\end{equation}
%
where we expressed the dispersion relation with the density of states $D(\varepsilon)$.
\subsection*{DMFT-loop}
Since the parameters of the effective model remain unknown, in order to solve the problem one uses an iterative procedure, in practice following the loop: 
\begin{enumerate}
\item Start with an initial guess for $G_0$.
\item Compute the AIM self-energy $\Sigma_{\text{imp}}$ by means of second order perturbation theory.
\item Compute the lattice problem local Green's function $G_{\text{loc}}$.
\item Update $G_0$ with imposing the self-consistency relation between the impurity Green's function and the local one, $G_{\text{loc}} \overset{!}{=} G$, therefore setting
  \begin{equation}
G_{0,\text{new}}^{-1} = G_{\text{loc}}^{-1} + \Sigma_{\text{imp}}.
\label{eq:updateG0}
  \end{equation}
\item Iterate till convergence.
\end{enumerate}
Finally, once the lattice local Green's function has been obtained for the set of values $\{i\omega_n\}$, we interpolate it using the Padé approximation, and eventually we are able to compute the Spectral Function via analytic continuation of the interpolation function.


